### 2022-11-25 intro.Rnw
paper 2 start
paper 2

\chapter*{Introduction}
\addcontentsline{toc}{chapter}{Introduction}

@@

\begin{table}[!htb]
  \centering
  \caption{Statistical criteria for assessing replication success which have
    been used in practice \citep{Opensc2015, Camerer2016, Camerer2018, Cova2018,
      Errington2021}.}
  \label{tab:rsdef}
  \begin{tabular}{p{0.21\textwidth}p{0.7\textwidth}}
    \toprule
    Criterion type & Description \\
    \midrule
    Significance & The original and replication $p$-values are smaller than
                   a threshold $\alpha$ and their effect estimates show
                   the same direction
                   ($p_{o} < \alpha$, $p_{r} < \alpha$, and
                   $\sign(\that_{o}) = \sign(\that_{r})$). Usually $\alpha = 5\%$.
    \\
                   & \\
    Meta-analytic significance & The meta-analytic $p$-value is smaller than a threshold
                                 $\alpha$ ($p_{m} = 2(1 - \Phi(|\that_{m}|/\sigma_{m})) < \alpha$)
                                 where $\that_{m} =
                                 (\that_{o}/\sigma_{o}^{2} + \that_{r}/\sigma^{2}_{r})\sigma^{2}_{m}$
                                 is the  pooled effect estimate with standard error
                                 $\sigma_{m} = (1/\sigma^{2}_{o} + 1/\sigma^{2}_{r})^{-1/2}$.
                                 Usually $\alpha = 5\%$. \\
                   & \\
    Relative effect size & The effect estimate of the replication study goes in the same
                           direction and is at least as large as $d_{\text{min}}$ times
                           the original one ($d = \that_{r}/\that_{o} \geq d_{\text{min}}$).
                           Usually $d_{\text{min}} = 1$. \\
                   & \\
    Confidence interval & The replication effect estimate is contained in the
                          $(1 - \alpha)$ original confidence interval
                          ($\that_{r} \in [\that_{o} \pm z_{\alpha/2}\sigma_{o}]$).
                          Sometimes, also defined as the original effect estimate is
                          contained in the $(1 - \alpha)$ replication confidence interval
                          ($\that_{o} \in [\that_{r} \pm z_{\alpha/2}\sigma_{r}]$).
                          Usually $\alpha = 5\%$. \\
                   & \\
    Prediction interval ($Q$-test) & The replication effect estimate is contained with its
                                     $(1 - \alpha)$ prediction interval
                                     ($\that_{r} \in [\that_{o} \pm z_{\alpha/2}\sqrt{\sigma_{o}^{2}
                                     + \sigma_{r}^{2}}]$). Usually $\alpha = 5\%$.
                                     This criterion is equivalent to $p_{Q} \geq \alpha$ where
                                     $p_{Q}$ is the $p$-value from the meta-analytic $Q$-test\footnotemark.\\
    \bottomrule
  \end{tabular}
\end{table}
\footnotetext{The $p$-value from the $Q$-test is
  $p_{Q} = 2\{1 - \Phi(\surd Q)\}$ where
  $Q = \sum_{i \in \{o, r\}} (\that_{i} - \that_{m})^{2}\sigma_{i}^{-2} = (\that_{o}- \that_{r})^{2}(\sigma^{2}_{o} + \sigma^{2}_{r})^{-1}$
  is the $Q$-statistic and $\that_{m}$ is the fixed-effects pooled estimate as
  defined in Table~\ref{tab:rsdef}.}

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion defines replication success by simultaneous statistical significance
of original and replication study along with their effect estimates showing the
same direction. This approach is also called the ``two-trials rule''
\citep{Senn2008} in drug regulation or ``vote-counting'' in meta-analysis
\citep{Cooper2019}. The criterion can similarly be defined through one-sided
$p$-values, so that the original and replication effect estimate are not
required to show the same direction as this is taken into account by the
$p$-values. Some replication projects \citep{Opensc2015, Errington2021} also
defined replication success via simultaneous non-significance of original and
replication study ($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this
definition ``replication success'' can almost always be achieved by conducting
original and replication study with just very few samples so that the $p$-values
are large. The approach is also logically questionable as it could be seen as an
instance of the ``absence of evidence fallacy'' \citep{Altman1995} meaning that
the failure to find evidence against the null hypothesis is erroneously
interpreted as evidence for the null hypothesis. Meta-analytic extensions of the
significance approach define replication success by significance of a combined
effect estimate\footnote{The meta-analytic approach can also be given a Bayesian
  interpretation: A flat prior distribution for the underlying effect size
  $\theta$ is updated by the data from original and replication study.
  Replication success defined by a meta-analytic $p$-value being smaller than
  $\alpha$ is then equivalent to replication success via a Bayesian posterior
  probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
  respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$
  depending on the direction of combined estimate.}. Typically, the assumption
of a common underlying effect size is seen as reasonable so that fixed-effects
meta-analysis is used for pooling. Random-effects meta-analysis has mostly been
used if more than one replication study of the same original study are conducted
since in this case replicators are often interested in also understanding
between-replication heterogeneity \citep[\eg{} in][]{Klein2018}.

@@

Based on these observations, Box proposed two general approaches for conducting
Bayesian model criticism. First, the predictive density of the observed data (or
the value of a ``checking function'' applied to the observed data) can be
compared to its reference distribution via a \emph{prior predictive $p$-value}
\begin{align}
  p_{\text{Box}}
  = \Pr\left\{f(\text{data} \given \text{M}) <
  f(\text{observed data} \given \text{M})\right\},
  \label{eq:pppval}
\end{align}
that is, the probability of obtaining data with lower predictive density (``more
surprising'' data) than the observed data. The lower the $p$-value
$p_{\text{Box}}$, the more incompatibility between the observed data and the
assumed model M. This approach was used by \citet{Held2020}, and it will soon
be explained in more detail.
% The literature on Bayesian model criticism following \citet{Box1980} also
% mostly focused on prior predictive $p$-values. However, what has been mostly
% ignored is that Box also mentioned an alternative approach.
However, Box also mentioned a second approach which has mostly been forgotten.
If a second ``benchmarking'' model $\text{M}_{2}$ alternative to the model under
investigation $\text{M}_{1}$ is available, Box proposed that the \emph{prior
  predictive ratio}
\begin{align*}
  \text{PR}_{\text{Box}}
  = \frac{f(\text{observed data} \given \text{M}_{1})}{f(\text{observed data} \given \text{M}_{2})},
\end{align*}
the ratio of predictive densities from the observed data under both models,
could be used to judge the relative adequacy of model $\text{M}_{1}$. Again, the
lower the predictive ratio $\text{PR}_{\text{Box}}$, the less compatible the
observed data with the model $\text{M}_{1}$. Bayesian model criticism approaches
based on predictive ratios will be used in later parts of this
thesis\footnote{Some readers may have noted that the predictive ratio is also
  the Bayes factor \citep{Jeffreys1961, Good1958} contrasting model
  $\text{M}_{1}$ to $\text{M}_{2}$. The prior predictive ratio model criticism
  approach is therefore particularly useful in combination with reverse-Bayes
  procedures based on Bayes factors, as will be demonstrated in this thesis.}.

>>>>

\chapter*{Introduction}
\phantomsection
\addcontentsline{toc}{chapter}{Introduction}

@@

\begin{table}[!htb]
  \centering
  \caption{Statistical criteria for assessing replication success which have
    been used in practice \citep{Opensc2015, Camerer2016, Camerer2018, Cova2018,
      Errington2021}.}
  \label{tab:rsdef}
  \begin{tabular}{p{0.21\textwidth}p{0.7\textwidth}}
    \toprule
    Criterion type & Description \\
    \midrule
    Significance & The original and replication $p$-values are smaller than
                   a threshold $\alpha$ and their effect estimates show
                   the same direction
                   ($p_{o} < \alpha$, $p_{r} < \alpha$, and
                   $\sign(\that_{o}) = \sign(\that_{r})$). Usually $\alpha = 5\%$.
    \\
                   & \\
    Meta-analytic significance & The meta-analytic $p$-value is smaller than a threshold
                                 $\alpha$ ($p_{m} = 2(1 - \Phi(|\that_{m}|/\sigma_{m})) < \alpha$)
                                 where $\that_{m} =
                                 (\that_{o}/\sigma_{o}^{2} + \that_{r}/\sigma^{2}_{r})\sigma^{2}_{m}$
                                 is the  pooled effect estimate with standard error
                                 $\sigma_{m} = (1/\sigma^{2}_{o} + 1/\sigma^{2}_{r})^{-1/2}$.
                                 Usually $\alpha = 5\%$. \\
                   & \\
    Relative effect size & The effect estimate of the replication study goes in the same
                           direction and is at least as large as $d_{\text{min}}$ times
                           the original one ($d = \that_{r}/\that_{o} \geq d_{\text{min}}$).
                           Usually $d_{\text{min}} = 1$. \\
                   & \\
    Confidence interval & The replication effect estimate is contained in the
                          $(1 - \alpha)$ original confidence interval
                          ($\that_{r} \in [\that_{o} \pm z_{\alpha/2}\sigma_{o}]$).
                          Sometimes, also defined as the original effect estimate is
                          contained in the $(1 - \alpha)$ replication confidence interval
                          ($\that_{o} \in [\that_{r} \pm z_{\alpha/2}\sigma_{r}]$).
                          Usually $\alpha = 5\%$. \\
                   & \\
    Prediction interval ($Q$-test) & The replication effect estimate is contained with its
                                     $(1 - \alpha)$ prediction interval
                                     ($\that_{r} \in [\that_{o} \pm z_{\alpha/2}\sqrt{\sigma_{o}^{2}
                                     + \sigma_{r}^{2}}]$). Usually $\alpha = 5\%$.
                                     This criterion is equivalent to $p_{Q} \geq \alpha$ where
                                     $p_{Q}$ is the $p$-value from the meta-analytic $Q$-test.\footnotemark\\
    \bottomrule
  \end{tabular}
\end{table}
\footnotetext{The $p$-value from the $Q$-test is
  $p_{Q} = 2\{1 - \Phi(\surd Q)\}$ where
  $Q = \sum_{i \in \{o, r\}} (\that_{i} - \that_{m})^{2}\sigma_{i}^{-2} = (\that_{o}- \that_{r})^{2}(\sigma^{2}_{o} + \sigma^{2}_{r})^{-1}$
  is the $Q$-statistic and $\that_{m}$ is the fixed-effects pooled estimate as
  defined in Table~\ref{tab:rsdef}.}

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion defines replication success by simultaneous statistical significance
of original and replication study along with their effect estimates showing the
same direction. This approach is also called the ``two-trials rule''
\citep{Senn2008} in drug regulation or ``vote-counting'' in meta-analysis
\citep{Cooper2019}. The criterion can similarly be defined through one-sided
$p$-values, so that the original and replication effect estimate are not
required to show the same direction as this is taken into account by the
$p$-values. Some replication projects \citep{Opensc2015, Errington2021} also
defined replication success via simultaneous non-significance of original and
replication study ($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this
definition ``replication success'' can almost always be achieved by conducting
original and replication study with just very few samples so that the $p$-values
are large. The approach is also logically questionable as it could be seen as an
instance of the ``absence of evidence fallacy'' \citep{Altman1995} meaning that
the failure to find evidence against the null hypothesis is erroneously
interpreted as evidence for the null hypothesis. Meta-analytic extensions of the
significance approach define replication success by significance of a combined
effect estimate.\footnote{The meta-analytic approach can also be given a
  Bayesian interpretation: A flat prior distribution for the underlying effect
  size $\theta$ is updated by the data from original and replication study.
  Replication success defined by a meta-analytic $p$-value being smaller than
  $\alpha$ is then equivalent to replication success via a Bayesian posterior
  probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
  respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$
  depending on the direction of combined estimate.} Typically, the assumption of
a common underlying effect size is seen as reasonable so that fixed-effects
meta-analysis is used for pooling. Random-effects meta-analysis has mostly been
used if more than one replication study of the same original study are conducted
since in this case replicators are often interested in also understanding
between-replication heterogeneity \citep[\eg{} in][]{Klein2018}.

@@

Based on these observations, Box proposed two general approaches for conducting
Bayesian model criticism. First, the predictive density of the observed data (or
the value of a ``checking function'' applied to the observed data) can be
compared to its reference distribution via a \emph{prior predictive $p$-value}
\begin{align}
  p_{\text{Box}}
  = \Pr\left\{f(\text{data} \given \text{M}) <
  f(\text{observed data} \given \text{M})\right\},
  \label{eq:pppval}
\end{align}
that is, the probability of obtaining data with lower predictive density (``more
surprising'' data) than the observed data. The lower the $p$-value
$p_{\text{Box}}$, the more incompatibility between the observed data and the
assumed model M. This approach was used by \citet{Held2020}, and it will soon
be explained in more detail.
% The literature on Bayesian model criticism following \citet{Box1980} also
% mostly focused on prior predictive $p$-values. However, what has been mostly
% ignored is that Box also mentioned an alternative approach.
However, Box also mentioned a second approach which has mostly been forgotten.
If a second ``benchmarking'' model $\text{M}_{2}$ alternative to the model under
investigation $\text{M}_{1}$ is available, Box proposed that the \emph{prior
  predictive ratio}
\begin{align*}
  \text{PR}_{\text{Box}}
  = \frac{f(\text{observed data} \given \text{M}_{1})}{f(\text{observed data} \given \text{M}_{2})},
\end{align*}
the ratio of predictive densities from the observed data under both models,
could be used to judge the relative adequacy of model $\text{M}_{1}$. Again, the
lower the predictive ratio $\text{PR}_{\text{Box}}$, the less compatible the
observed data with the model $\text{M}_{1}$. Bayesian model criticism approaches
based on predictive ratios will be used in later parts of this
thesis.\footnote{Some readers may have noted that the predictive ratio is also
  the Bayes factor \citep{Jeffreys1961, Good1958} contrasting model
  $\text{M}_{1}$ to $\text{M}_{2}$. The prior predictive ratio model criticism
  approach is therefore particularly useful in combination with reverse-Bayes
  procedures based on Bayes factors, as will be demonstrated in this thesis.}
